---
title: "Synthesize Dataset"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output: html_document
---

Load the libraries
```{r}
library(tidyverse)
library(raster)
library(sf)
library(sp)
library(dplyr)
library(tidyr)
library(tmap)
library(lubridate)
```


Read all the data sets
```{r}

#setwd("/capstone/kelpgeomod/")

# Nutrients
interp_phosphate <- brick("/capstone/kelpgeomod/analysis_data/inteprolate_idw/phosphate_quarterly.tif")
interp_nitrate_nitrite <- brick("/capstone/kelpgeomod/analysis_data/inteprolate_idw/nitrate_nitrite_quarterly.tif")
interp_ammonium <- brick("/capstone/kelpgeomod/analysis_data/inteprolate_idw/ammonium_quarterly.tif")
joined_nutrients <- readRDS("/capstone/kelpgeomod/intermediate_data/joined_nutrients/joined_nutrients.RDS")

# Sea Surface Temperature
sst <- brick("/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_raster.tif")

# Depth
depth <- raster("/capstone/kelpgeomod/intermediate_data/depth/SBC_depth.tif")

#Kelp
kelp_biomass <- brick("/capstone/kelpgeomod/intermediate_data/area_biomass/biomass_brick_standardized.tif")
kelp_area <- brick("/capstone/kelpgeomod/intermediate_data/area_biomass/area_brick_standardized.tif")

# Mask
mask <- raster("/capstone/kelpgeomod/intermediate_data/mask/mask_rast.tif")


```

# DEPTH
```{r}
# Convert the Raster brick to a data frame
depth_df <- as.data.frame(depth, xy = TRUE, na.rm = TRUE)

# Define the years and quarters to be added
years <- 2014:2022
quarters <- 1:4

# Create an empty data frame to store the results
depth_year_quarter_df <- data.frame()

# Loop over quarters and years
for (q in quarters) {
  for (y in years) {
    # Add a row to the data frame with the current year and quarter
    row <- cbind(depth_df, year = y, quarter = q)
    depth_year_quarter_df <- rbind(depth_year_quarter_df, row)
  }
}


# Create a data frame for the Time Series
depth_year_quarter_df <- depth_year_quarter_df %>% 
  rename(lon = x, lat = y, depth = depth_long) %>% 
  dplyr::select(year, quarter, lon, lat, depth)
  
plot(depth)
```

# SST
```{r}
# Re-add names since they are lost when you read back in
years <- c(2014:2022)
quarters <- c(1:4)

year_quarter <- c()
for (i in seq_along(years)) {
  for (j in seq_along(quarters)) {
  year_quarter_new <- paste0("layer", years[i], "_", quarters[j])
  year_quarter <- append(year_quarter, year_quarter_new)
  }
}

names(sst) <- paste(year_quarter)

# Check the extent and resolution of the raster object
identical(extent(sst), extent(mask))
identical(res(sst), res(mask))
identical(crs(sst), crs(mask))
identical(origin(sst), origin(mask))

# Convert the raster brick to a data frame
sst_year_quarter_df <- as.data.frame(sst, xy = TRUE, na.rm = TRUE)
sst_year_quarter_df <- sst_year_quarter_df %>%
  pivot_longer(cols = starts_with("l"), # Select all columns that start with "l" for "layer"
               names_to = "year_quarter", # Name of the new column to be created for year_quarter
               values_to = "sst") %>%     # Name of the new column to be created for sst values
  separate(year_quarter, c("year", "quarter")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(year = str_remove(year, "layer")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and lon
  dplyr::select(year, quarter, lon, lat, sst)  # Reorder the columns as needed

```


# KELP
## Kelp Area
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(kelp_area), extent(mask))
identical(res(kelp_area), res(mask))
identical(crs(kelp_area), crs(mask))
identical(origin(kelp_area), origin(mask))

#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(kelp_area)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(kelp_area) <- new_names


#Create a Data Frame from this list
# Convert the Raster brick to a data frame
kelp_area_year_quarter_df <- as.data.frame(kelp_area, xy = TRUE, na.rm = TRUE)
kelp_area_year_quarter_df <- kelp_area_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "Q"
               names_to = "quarter_year", # Name of the new column to be created for quarter-year
               values_to = "area") %>%     # Name of the new column to be created for SST values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the quarter_year column into quarter and year columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the Q of the quarter values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  dplyr::select(year, quarter, lon ,lat, area)  # Reorder the columns as needed


```

## Kelp Biomass
```{r}
# Re-add names since they are lost when you read back in
years <- c(2014:2022)
quarters <- c(1:4)

year_quarter <- c()
for (i in seq_along(years)) {
  for (j in seq_along(quarters)) {
  year_quarter_new <- paste0("layer", years[i], "_", quarters[j])
  year_quarter <- append(year_quarter, year_quarter_new)
  }
}

names(kelp_biomass) <- paste(year_quarter)

# Check the extent and resolution of the raster object
identical(extent(kelp_biomass), extent(mask))
identical(res(kelp_biomass), res(mask))
identical(crs(kelp_biomass), crs(mask))
identical(origin(kelp_biomass), origin(mask))

# Convert the raster brick to a data frame
kelp_biomass_year_quarter_df <- as.data.frame(kelp_biomass, xy = TRUE, na.rm = FALSE)
kelp_biomass_year_quarter_df <- kelp_biomass_year_quarter_df %>%
  pivot_longer(cols = starts_with("l"), # Select all columns that start with "l" for "layer"
               names_to = "year_quarter", # Name of the new column to be created for year_quarter
               values_to = "biomass") %>%     # Name of the new column to be created for kelp values
  separate(year_quarter, c("year", "quarter")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(year = str_remove(year, "layer")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  dplyr::select(year, quarter, lon, lat, biomass)  # Reorder the columns as needed

```

# NUTRIENTS
## Joined Nutrients
```{r}
# Extract the coordinates from the geometry column
coordinates <- st_coordinates(joined_nutrients)

# Convert the coordinates to a data frame
coordinates_df <- as.data.frame(coordinates)

# Rename the columns to lon and lat
names(coordinates_df) <- c("lon", "lat")

# Add the lon and lat columns to the original sf object
joined_nutrients <- cbind(joined_nutrients, coordinates_df) %>% 
  st_drop_geometry(joined_nutrients, "geometry") %>% 
  dplyr::select("quarter", "year", "lon", "lat", "source", "nitrate_nitrite" ,"ammonium", "phosphate")

```

## Interpolated Nutrients
### Phosphate
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_phosphate), extent(mask))
identical(res(interp_phosphate), res(mask))
identical(crs(interp_phosphate), crs(mask))
identical(origin(interp_phosphate), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
phosphate_df <- as.data.frame(interp_phosphate, xy = TRUE, na.rm = TRUE)

names(phosphate_df)[3:6] <- c("1", "2", "3", "4")

phosphate_seasonal <- phosphate_df %>%
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "phosphate") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(phosphate = mean(phosphate, na.rm = T))
  
  
```

### Nitrate_Nitrite
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_nitrate_nitrite), extent(mask))
identical(res(interp_nitrate_nitrite), res(mask))
identical(crs(interp_nitrate_nitrite), crs(mask))
identical(origin(interp_nitrate_nitrite), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
nitrate_nitrite_df <- as.data.frame(interp_nitrate_nitrite, xy = TRUE, na.rm = TRUE)
nitrate_nitrite_seasonal <- nitrate_nitrite_df %>%
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "nitrate_nitrite") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(nitrate_nitrite = mean(nitrate_nitrite, na.rm = T))
```

### Ammonium
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_ammonium), extent(mask))
identical(res(interp_ammonium), res(mask))
identical(crs(interp_ammonium), crs(mask))
identical(origin(interp_ammonium), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
ammonium_df <- as.data.frame(interp_ammonium, xy = TRUE, na.rm = TRUE)
ammonium_seasonal <- ammonium_df %>%
  rename(lon = "x", lat = "y")  %>%  # Rename the x and y columns to lat and long
  pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "ammonium") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(ammonium = mean(ammonium, na.rm = T))

```

# Time Series Visualization
## Normalize data sets
```{r}
# Compute time series wrangling for each year_quarter data frames
#DEPTH
depth_time_series <- depth_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(depth = mean(depth)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# SST
sst_time_series <- sst_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# KELP
# Kelp Area
kelp_area_time_series <- kelp_area_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(area = sum(area, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# Kelp Biomass
kelp_biomass_time_series <- kelp_biomass_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(biomass = sum(biomass, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# NUTRIENTS
nutrients_time_series <- joined_nutrients %>% 
  group_by(year, quarter) %>% 
  summarise(nitrate = mean(nitrate, na.rm = T),
            nitrite = mean(nitrite, na.rm = T),
            nitrate_nitrite = mean(nitrate_nitrite, na.rm = T),
            ammonium = mean(ammonium, na.rm = T),
            phosphate = mean(phosphate, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))
```

## Join
```{r}
# Merge the all normalized data sets
time_series_merge <- depth_time_series %>% 
  left_join(sst_time_series, by = c("year", "quarter")) %>% 
  left_join(kelp_area_time_series, by = c("year", "quarter")) %>% 
  left_join(kelp_biomass_time_series, by = c("year", "quarter")) %>% 
  left_join(nutrients_time_series, by = c("year", "quarter"))

# Pivot the data into long format
time_series_long <- pivot_longer(time_series_merge, 
                                 cols = 3:11, 
                                 names_to = "variable", 
                                 values_to = "value") %>% 
   mutate(date = as.Date(paste(year, as.integer(quarter)*3, "01", sep = "-")))
```

## Set the variable colors
```{r}
my_colors <- c("depth" = "#2E2E2E", 
               "sst" = "#FD8D3C", 
               "area" = "#3CB44B", 
               "biomass" = "#91CF60", 
               "nitrate" = "#1F78B4", 
               "nitrite" = "#A6CEE3",
               "nitrate_nitrite" = "#984EA3", 
               "ammonium" = "#E41A1C", 
               "phosphate" = "#FF7F00")

```

## Facet Wrap Visualization
```{r}
ggplot(time_series_long, aes(x = date, y = value, color = variable)) +
  geom_line() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_color_manual(values = my_colors) +
  labs(x = "Date", y = "Value", color = "Variable", title = "Oceanographic Time Series") +
  facet_wrap(~ variable, scales = "free_y") +
  theme_minimal()

```

## Individual Visualization
```{r}

# Define the color palette
my_colors <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02", "#a6761d", "#666666", "#bdbdbd")

# Create a list of unique variable names
variable_names <- unique(time_series_long$variable)

# Create a for loop to create a separate ggplot for each variable
for (i in seq_along(variable_names)) {
  # Subset the data for the current variable
  data_subset <- subset(time_series_long, variable == variable_names[i])
  
  # Create a ggplot for the current variable
  p <- ggplot(data_subset, aes(x = date, y = value, color = variable_names[i])) +
    geom_line(color = my_colors[i]) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(x = "Date", 
         y = variable_names[i], 
         title = "kelpGeoMod Time Series",
         subtitle = paste("Oceanographic factor:", variable_names[i])) +
    theme_minimal()
  
  # Print the ggplot
  print(p)
}



```

# Seasonal Rasters
## Normalize data sets
```{r}
# Compute wrangling for each data frame to get the seasonal mean
#DEPTH
depth_seasonal <- depth_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(depth = mean(depth)) %>% 
  mutate(quarter = as.factor(quarter))

#SST
sst_seasonal <- sst_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(quarter = as.factor(quarter))

#KELP
# Kelp Area
kelp_area_seasonal <- kelp_area_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(area = mean(area, na.rm = T)) %>% 
  mutate(quarter = as.factor(quarter))

# Kelp Biomass
kelp_biomass_seasonal <- kelp_biomass_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(biomass = mean(biomass, na.rm = T)) %>% 
  mutate(quarter = as.factor(quarter))

```

## Seasonal Join
```{r}
# Merge the all normalized seasonal data sets
seasonal_merge <- depth_seasonal %>% 
  left_join(sst_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(kelp_area_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(kelp_biomass_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(phosphate_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(nitrate_nitrite_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(ammonium_seasonal, by = c("quarter", "lon", "lat"))
  
```


Save the 36 raster brick
```{r}
# Write RasterBrick Quarter object to file
write_csv(seasonal_merge, "/capstone/kelpgeomod/analysis_data/synthesized/interpolated_join.csv")

write_csv(time_series_merge, 
            path = "/capstone/kelpgeomod/analysis_data/synthesized/time_series_join.csv")
```
