---
title: "SST"
author: "Javier Patron"
date: "r Sys.Date()"
---

# Sea Surface Temperature

The Sea Surface Temperature (SST) is in the form of a data set called "MUR-JPL-L4-GLOB-v4.1." This data set is produced by the Jet Propulsion Laboratory (JPL) and is part of the Multi-scale Ultra-high Resolution (MUR) sea surface temperature project. The metadata for this data set provides information about the characteristics and properties of the SST data.

## Reading and Loading the SST Data

Load the libraries
```{r}
library(tidyverse)
library(httr)
library(ncdf4)
library(tmap)
library(raster)
library(tictoc)
library(terra)
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(gganimate)
library(animation)
library(sf)
```

Read in the mask
```{r}
mask <- raster("/capstone/kelpgeomod/intermediate_data/mask/mask_rast.tif")
```

Read one image for testing
```{r}

# Open the first file from the folder
file <- nc_open("/capstone/kelpgeomod/raw_data/SST/GRHSST_lvl4/20230101090000-JPL-L4_GHRSST-SSTfnd-MUR-GLOB-v02.0-fv04.1.dap.nc4")

# Retrieve the "analysed_sst" variable from a NetCDF file, which contains sea surface temperature data, and assigns it to the "sst" variable for further analysis or manipulation.
sst <- ncvar_get(file, "analysed_sst")

# Convert Kelvin to Celsius
sst <- sst - 273.15  

# Convert to raster for visualization
sst_rast <- raster(sst) 

# Flip the raster by 90 degrees to the left
flipped_data <- t(flip(sst_rast, 1))

plot(flipped_data)

```

# General Info of the raw image
```{r}
plot(sst_rast) #Plot the raw image
crs(sst_rast)
res(sst_rast)
plot(flipped_data)

```



#### *Insights for netCDF files:* NetCDF files are self-describing, meaning that they include metadata that describes the contents of the file, such as variable names, units, and dimensions. This makes it easy to understand the data without having to consult external documentation. netCDF is popular becuase it can store various data types and numerical data types and can be read and written on various operating systems.


Check the image info and use the mask
```{r}
#USING THE RESAMPLE
# Set the extent of the raster to match the mask extent
raster_data <- setExtent(flipped_data, extent(mask))

# Resample the raster to match the mask resolution
resampled_data <- resample(raster_data, mask, method = "bilinear")

# Apply the mask to the resampled raster
sst_masked <- mask(resampled_data, mask)

#Check dimentions
dim(sst_masked)

plot(sst_masked)

```

# Creating a For Loop for the quarterly raster stack

Set up the directory path where the files are located:
```{r}
data_dir <- "/capstone/kelpgeomod/raw_data/SST/GRHSST_lvl4"
```

Create a list of all the files in the directory:
```{r}
file_list <- list.files(data_dir, full.names=TRUE)
```

Create a Data Frame that contains the information from the files names and creates a new column in a data frame:
```{r}
#Filter for the desired dates
start_date <- as.Date("2014-01-01")
end_date <- as.Date("2022-12-31")

file_df <- data.frame(file_name=file_list) %>%
  mutate(date = as.Date(str_extract(file_name, "\\d{8}"), 
                        format="%Y%m%d")) %>% 
  filter(date >= start_date & date <= end_date)

```

Create a function that reads in each file and convert it to a raster:
```{r}

ncdf_to_raster <- function(nc_file) {
  file <- nc_open(nc_file) # Read the 
  sst <- ncvar_get(file, "analysed_sst")
  sst <- sst - 273.15  # Convert Kelvin to Celsius
  sst_rast <- raster(sst)
  
  # Flip the raster by 90 degrees to the left
  flipped_data <- t(flip(sst_rast, 1))
  
  # Set the extent of the raster to match the mask extent
  raster_data <- setExtent(flipped_data, extent(mask))
  
  # Resample the raster to match the mask resolution
  resampled_data <- resample(raster_data, mask, method="bilinear")
  
  # Apply the mask to the resampled raster
  sst_masked <- mask(resampled_data, mask)
  
  nc_close(file) # Close the netCDF file
  
  return(sst_masked)
}

```

Do a test for January 2014
```{r}

# Subset file_df to include only files from January 2014
jan_2014_files <- file_df %>% 
  filter(date < '2014-02-01')

# Use lapply() to create a list of rasters for January 2014
raster_list <- lapply(jan_2014_files$file_name, ncdf_to_raster)

# Convert to RasterStack object
jan_raster_stack <- stack(raster_list)

# Calculate the mean of all layers
jan_raster_mean <- calc(jan_raster_stack, mean)

plot(jan_raster_mean)

```

## Normalizing the SST files

a)  Create a function that builds a Data Frame with the year and month we are interested on
b)  Apply the ncdf_to_raster function that reads in the image, converts to Celsius and transforms each image into a raster file

```{r}

create_raster_stack <- function(year, month) {
  
  # Subset file_df to include only files from the given year and month
  files <- file_df %>% 
    filter(year(date) == year, month(date) == month)
  
  # Use lapply() to create a list of rasters for the given year and month
  raster_list <- lapply(files$file_name, ncdf_to_raster)
  
  # Convert to RasterStack object
  raster_stack <- stack(raster_list)
  
  # Calculate the mean of all layers
  raster_mean <- calc(raster_stack, mean)
  
  return(raster_mean)
}

```

Test the function we just did for August 2017

```{r}
tic()
create_raster_stack(2017, 8)
toc()
```


## Now lets create the raster brick *per quarter*

Create a for loop that:
a)  Uses the function `create_raster_stack` to build a raster mean per quarter
b)  Appends that new raster to a list
c)  Executes the mean per quarter
d)  Stacks all quarter asters into a brick stack

```{r}
# Step a)
create_raster_stack_quarterly <- function(year, quarter) {
  
  # Define the start and end months for the given quarter
  start_quarter <- (quarter - 1) * 3 + 1
  end_quarter <- quarter * 3
  
  # Subset file_df to include only files from the given year and quarter
  files <- file_df %>% 
    filter(year(date) == year & month(date) >= start_quarter & month(date) <= end_quarter)
  
  # Use lapply() to create a list of rasters for the given year and quarter
  raster_list <- lapply(files$file_name, ncdf_to_raster)
  
  # Convert to RasterStack object
  raster_stack <- stack(raster_list)
  
  # Calculate the mean of all layers
  raster_mean <- calc(raster_stack, mean)
  
  return(raster_mean)
  
}
```

Test and visualize the function with a example
```{r}
test_spring15 <- create_raster_stack_quarterly(2015, 2)
plot(test_spring15)

```


```{r}

# Step b. Create empty list to store raster stacks
raster_stack_list <- list()

# Step c.Executes the for loop by using the function and calculates the mean of SST per quarter
tic()
# Loop through all quarters from Q1 2014 to Q4 2022
for (year in 2014:2022) {
  for (quarter in 1:4) {
    
    # Create raster stack for current year and quarter
    raster_stack <- create_raster_stack_quarterly(year, quarter)
    
    # Add raster stack to list
    raster_stack_list[[length(raster_stack_list) + 1]] <- raster_stack
    
  }
}


toc()

# Convert list of raster stacks to RasterBrick object
raster_quarter_brick <- brick(raster_stack_list)

# Check the object that we just created
class(raster_quarter_brick)
names(raster_quarter_brick)
sum(length(names(raster_quarter_brick)))

```

Create the names for the raster files
```{r}
# Get current names of layers in raster_brick
current_names <- names(raster_quarter_brick)

# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")

# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))

# Rename layers in raster_brick with new names
names(raster_quarter_brick) <- new_names

```

Save the 36 raster brick
```{r}
# Write RasterBrick Quarter object to file
writeRaster(raster_quarter_brick, 
            filename = "/capstone/kelpgeomod/intermediate_data/sst/sst_all_quarter_raster", 
            format = "GTiff", 
            overwrite=TRUE)
```

Create the raster brick per season
```{r}
calc_seasonal_means_brick <- function(rast_to_convert) {

  quarter_sets <- list(seq(from = 1, to = 36, by = 4), # Q1s (winter)
                       seq(from = 2, to = 36, by = 4), # Q2s (spring)
                       seq(from = 3, to = 36, by = 4), # Q3s (summer)
                       seq(from = 4, to = 36, by = 4)) # Q4s (fall)
  
  all_seasons_brick <- brick() # set up brick to hold averaged layers for each season (will have 4 layers at the end)
  
  for (i in seq_along(quarter_sets)) {
    
    season_brick_holder <- brick() # hold all layers for one season, then reset for next season
    
    for (j in quarter_sets[[i]]) {
      season_brick <- brick() # hold single layer in a season, then reset for next layer
      season_brick <- addLayer(season_brick, rast_to_convert[[j]]) # add single layer to initialized brick
      season_brick_holder <- addLayer(season_brick_holder, season_brick) # add this layer to the holder for this season, and repeat until have all layers from season
    }

    season_averaged_layer <- calc(season_brick_holder, mean) # after having all layers from season, take the mean
    all_seasons_brick <- addLayer(all_seasons_brick, season_averaged_layer) # add mean to the brick holding all averaged layers, and then repeat for the next season
}

writeRaster(all_seasons_brick[[1]], "/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_1.tif", format = "GTiff")
writeRaster(all_seasons_brick[[2]], "/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_2.tif", format = "GTiff")
writeRaster(all_seasons_brick[[3]], "/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_3.tif", format = "GTiff")
writeRaster(all_seasons_brick[[4]], "/capstone/kelpgeomod/intermediate_data/sst/sst_quarter_4.tif", format = "GTiff")
}

```


```{r}
calc_seasonal_means_brick(raster_quarter_brick)
```

Save the 4 raster brick
```{r}
# Write RasterBrick Quarter object to file
writeRaster(seasonal_sst_brick, 
            filename = "/capstone/kelpgeomod/intermediate_data/sst/sst_seasonal", 
            format = "GTiff",
            overwrite= TRUE)

```

