---
editor_options: 
  markdown: 
    wrap: 72
---

# Wallace Session 2023-05-29

Please find below the R code history from your *Wallace* v2.0.5 session.

You can reproduce your session results by running this R Markdown file
in RStudio.

Each code block is called a "chunk", and you can run them either
one-by-one or all at once by choosing an option in the "Run" menu at the
top-right corner of the "Source" pane in RStudio.

For more detailed information see <http://rmarkdown.rstudio.com>).

### Package installation

Wallace uses the following R packages that must be installed and loaded
before starting.

```{r}
library(spocc)
library(spThin)
library(dismo)
library(rgeos)
library(ENMeval)
library(wallace)
library(tidyverse)
```

The *Wallace* session code .Rmd file is composed of a chain of module
functions that are internal to *Wallace*. Each of these functions
corresponds to a single module that the user ran during the session. To
see the internal code for these module functions, click on the links in
the .Rmd file. Users are encouraged to write custom code in the .Rmd
directly to modify their analysis, and even modify the module function
code to further customize. To see the source code for any module
function, just type its name into the R console and press Return.

```{r}
# example:
# just type the function name and press Return to see its source code
# paste this code into a new script to edit it
occs_queryDb
```

Your analyses are below.

------------------------------------------------------------------------

## Analysis for *Macrocystis pyrifera* (Mp)

User CSV path with occurrence data. If the CSV file is not in the
current workspace, change to the correct file path
(e.g./Users/darwin/Documents/occs/").

```{r}
# NOTE: provide the folder path of the .csv file
occs_path <- "/capstone/kelpgeomod/new_file_structure/03-analysis-data/04-maxent-analysis/quarter_4"
occs_path <- file.path(occs_path, "kelp_presence_4.csv")
# get a list of species occurrence data
userOccs_Mp <- occs_userOccs(
  txtPath = occs_path, 
  txtName = "kelp_presence_4.csv", 
  txtSep = ",", 
  txtDec = ".")
occs_Mp <- userOccs_Mp$Macrocystis_pyrifera$cleaned
```

### Obtain environmental data

Using user-specified variables.

```{r}
## Specify the directory with the environmental variables
dir_envs_Mp <- "/capstone/kelpgeomod/new_file_structure/03-analysis-data/04-maxent-analysis/quarter_4"
envs_path <- file.path(dir_envs_Mp, c('depth_4.tif', 'nitrate_nitrite_4.tif', 'phosphate_4.tif'))
# Create environmental object 
envs_Mp <- envs_userEnvs(
  rasPath = envs_path,
  rasName = c('depth_4.tif', 'nitrate_nitrite_4.tif', 'phosphate_4.tif'),
  doBrick = FALSE)
occs_xy_Mp <- occs_Mp[c('longitude', 'latitude')]
occs_vals_Mp <- as.data.frame(raster::extract(envs_Mp, occs_xy_Mp, cellnumbers = TRUE))
# Remove duplicated same cell values
occs_Mp <- occs_Mp[!duplicated(occs_vals_Mp[, 1]), ]
occs_vals_Mp <- occs_vals_Mp[!duplicated(occs_vals_Mp[, 1]), -1]
# remove occurrence records with NA environmental values
occs_Mp <- occs_Mp[!(rowSums(is.na(occs_vals_Mp)) >= 1), ]
# also remove variable value rows with NA environmental values
occs_vals_Mp <- na.omit(occs_vals_Mp)
# add columns for env variable values for each occurrence record
occs_Mp <- cbind(occs_Mp, occs_vals_Mp)
```

### Process environmental data

Sampling of 800 background points and corresponding environmental data
using a user provided background extent with a 0 degree buffer.

```{r}
# Set the background to our area of environmental data plus a slight buffer
bgExt_Mp <- penvs_bgExtent(occs_Mp, bgSel = 'bounding box', bgBuf = 0.5)

bgMask_Mp <- penvs_bgMask(
  occs = occs_Mp,
  envs = envs_Mp,
  bgExt = bgExt_Mp)
# Sample background points from the provided area
bgSample_Mp <- penvs_bgSample(
  occs = occs_Mp,
  bgMask =  bgMask_Mp,
  bgPtsNum = 800)
# Extract values of environmental layers for each background point
bgEnvsVals_Mp <- as.data.frame(raster::extract(bgMask_Mp,  bgSample_Mp))
##Add extracted values to background points table
bgEnvsVals_Mp <- cbind(scientific_name = paste0("bg_", "Macrocystis pyrifera"), bgSample_Mp,
                            occID = NA, year = NA, institution_code = NA, country = NA,
                            state_province = NA, locality = NA, elevation = NA,
                            record_type = NA, bgEnvsVals_Mp)
```

### Partition occurrence data

Partition occurrences and background points for model training and
validation using "hierarchical checkerboard", a spatial partition method
with an aggregation factor of 2.

```{r}
# R code to get partitioned data
groups_Mp <- part_partitionOccs(
  occs = occs_Mp ,
  bg =  bgSample_Mp, 
  method = "cb2",
  bgMask = bgMask_Mp,
  aggFact = 2) 
```

### Build and Evaluate Niche Model

Generating a species distribution model using the maxent.jar algorithm
as implemented in ENMeval V2.0 (with clamping = FALSE). For tuning using
L, LQ, H, LQH, LQHP feature classes and regularization multipliers in
the 0.5, 4.5 range increasing by 0.2. Not using any categorical
predictor variables.

If this returns an error with rJava, please refer to "Setting Up Java
Version of Maxent" in the [Wallace help
Documentation](https://wallaceecomod.github.io/wallace/articles/tutorial-v2.html#obtain-environmental-data)

```{r}
set.seed(123)

# Run maxent model for the selected species using maxent
# model_Mp <- model_maxent(
#   occs = occs_Mp,
#   bg = bgEnvsVals_Mp,
#   user.grp = groups_Mp, 
#   bgMsk = bgMask_Mp,
#   rms = c(0.5, 4.5), 
#   rmsStep =  0.2,
#   fcs = c('L', 'LQ', 'H', 'LQH', 'LQHP'),
#   clampSel = FALSE,
#   algMaxent = "maxnet",
#   parallel = TRUE,
#   numCores = 6)


#--------------------------------maxent.jar strategy------------------------------

# Run maxent model for the selected species using maxent.jar 
# This needs rjava to run -- without rjava run above
# This gives access to feature importance, above does not
model_Mp <- model_maxent(
  occs = occs_Mp,
  bg = bgEnvsVals_Mp,
  user.grp = groups_Mp, 
  bgMsk = bgMask_Mp,
  rms = c(0.5, 4.5), 
  rmsStep =  0.2,
  fcs = c('L', 'LQ', 'H', 'LQH', 'LQHP'),
  clampSel = FALSE,
  algMaxent = "maxent.jar",
  parallel = TRUE,
  numCores = 30) # Change to realistic cores for your machine
```

### Model eval & selection

```{r}
# Make the results a dataframe
mod_res <- model_Mp@results

# Sort data frame by AICc column in ascending order
mod_res_sorted <- mod_res[order(mod_res$AICc), ]

# Select the row with the lowest AICc value
lowest_AIC_row <- mod_res_sorted[1, ]

# Retrieve the value of the "tune.args" column from the row with the lowest AICc
lowest_AIC_tune_args <- lowest_AIC_row$tune.args
lowest_AIC_tune_args # Use this in the following section
```

### Visualize

Generate a map of the Maxent generated model with no threshold

```{r}
# Select current model and obtain raster prediction
m_Mp <- model_Mp@models[["fc.LQ_rm.0.5"]] # Pulled from lowest_AIC_tune_args
predSel_Mp <- dismo::predict(
  m_Mp, bgMask_Mp,
  args = c(paste0("outputformat=", "cloglog"), 
           paste0("doclamp=", tolower(as.character(FALSE)))), 
  na.rm = TRUE)
#Get values of prediction
mapPredVals_Mp <- getRasterVals(predSel_Mp, "cloglog")
#Define colors and legend  
rasCols <- c("#2c7bb6", "#abd9e9", "#ffffbf", "#fdae61", "#d7191c")
legendPal <- colorNumeric(rev(rasCols), mapPredVals_Mp, na.color = 'transparent')
rasPal <- colorNumeric(rasCols, mapPredVals_Mp, na.color = 'transparent')
#Generate map
m <- leaflet() %>% addProviderTiles(providers$Esri.WorldTopoMap) 
m  %>%
  leaflet::addLegend("bottomright", pal = legendPal,
            title = "Predicted Suitability<br>(Training)",
            values = mapPredVals_Mp, layerId = "train",
            labFormat = reverseLabel(2, reverse_order = TRUE)) %>% 
  #add occurrence data
  addCircleMarkers(data = occs_Mp, lat = ~latitude, lng = ~longitude,
                   radius = 5, color = 'red', fill = TRUE, fillColor = "red",
                   fillOpacity = 0.2, weight = 2, popup = ~pop) %>% 
  ##Add model prediction
  addRasterImage(predSel_Mp, colors = rasPal, opacity = 0.7,
                 group = 'vis', layerId = 'mapPred', method = "ngb") %>%
 ##add background polygons
  addPolygons(data = bgExt_Mp,fill = FALSE,
              weight = 4, color = "blue", group = 'proj')
```

### Visualize

Generate a Maxent evaluation plot using "delta.AICc" as evaluation
statistic.

```{r}
# Generate an evaluation plot
maxentEvalPlot_Mp <- ENMeval::evalplot.stats(
   model_Mp,
   "delta.AICc",
   "rm",
   "fc")
#plot
maxentEvalPlot_Mp
```

### Variable Importance

```{r}
vars <- eval.variable.importance(model_Mp)[["fc.LQ_rm.0.5"]]
kable(vars)
```

### Data Plots

```{r}
# Depth
depth_dat <- cbind(as.data.frame(envs_Mp[[1]]), as.data.frame(predSel_Mp))
ggplot(depth_dat, aes(depth_4, layer)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Depth",
       y = "Predicted Habitat Suitability")

# Nitrogen
nitr_dat <- cbind(as.data.frame(envs_Mp[[2]]), as.data.frame(predSel_Mp))
ggplot(nitr_dat, aes(nitrate_nitrite_4, layer)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Nitrogen",
       y = "Predicted Habitat Suitability")

# Phosphate
phos_dat <- cbind(as.data.frame(envs_Mp[[3]]), as.data.frame(predSel_Mp))
ggplot(phos_dat, aes(phosphate_4, layer)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Phosphate",
       y = "Predicted Habitat Suitability")
```

```{r}
terra::writeRaster(predSel_Mp, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/04-maxent-analysis/results/maxent-quarter-4-output.tif", overwrite = TRUE)
```