---
title: "Untitled"
author: "Jessica French"
date: "`r Sys.Date()`"
output: html_document
---
# Purpose

This markdown combines the shape files from the USGS into a single raster that has the same classification scheme, extent, and resolution as the USGS coast files and the habitat mask. 

## Set up 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r label = "read in libraries"}
library(terra)
library(tmap)
library(sf)
library(tidyverse)

```

```{r label = "define data directory"}
# define the data dorectory
data_dir <- "/capstone/kelpgeomod/google_drive_download"
```

These files are in a slightly different format and have different information than the main USGS files. From the metadata the bottom classification is in the BOTTOM_ID column and is coded as h for hard bottom, m for mixed hard and s for sediment soft bottom. There is not additional information for what may be an anthropogenic feature so they cannot be removed from this data. 


```{r label = "define function for read in in files"}
# The following function reads in a shape file using st_read, transforms the crs to WGS 84 and selects the columns of interest.

USGS_channel_fun <- function(shape_file) {
  st_read(shape_file) |> 
  st_transform("EPSG:4326") |> 
  dplyr::select(c(BOTTOM_ID, geometry))  
}

```
    
```{r label = "read in files"}
# This entire code chunk is courtesy of ChatGPT.
# get list of subdirectories in working directory
subdirs <- list.dirs(file.path(data_dir, "01-raw-data/08-substrate-usgs-seafloormapping-raw/03-usgs-channel-islands"))

# Create empty list
USGS_islands <- list()

# Loop over subdirectories and read shapefiles
for (subdir in subdirs) {
  # Get list of shapefiles in current subdirectory
  shapefiles <- list.files(path = subdir, pattern = "\\.shp$", full.names = TRUE) 
  
  # Loop over shapefiles in current subdirectory and read them
  for (shapefile in shapefiles) {
    # Read shapefile using sf::st_read function and assign to a new object with the name of the shapefile
    shp <- USGS_channel_fun(shapefile)
    shp_name <- sub('\\.shp$', '', basename(shapefile))
    assign(shp_name, shp)
    
    # Append the named object to the list
    USGS_islands[[shp_name]] <- get(shp_name)
  }
}
```
    
```{r label = "combine into one data frame"}
# Combine list if data frames into one using bind rows. 
USGS_islands <-  bind_rows(USGS_islands) 

#plot(USGS_channel_islands$geometry)
```

```{r label = "read in mask"}
# read in mask

substrate_mask <- rast(file.path(data_dir, "02-intermediate-data/03-mask-intermediate/habitat-mask.tif"))
```

```{r label = "rasterize"}

usgs_islands_raster <- rasterize(USGS_islands, substrate_mask, field = "BOTTOM_ID") 

# set the names of the layer to be consistent with the other substrate files
terra::set.names(usgs_islands_raster, "substrate_class")

plot(usgs_islands_raster)

```

```{r label = "write to shape file "}
# An aux.xml metadata file is automatically written with each file. 
terra::writeRaster(usgs_islands_raster, "/capstone/kelpgeomod/google_drive_download/02-intermediate-data/08-substrate-intermediate/usgs-island-substrate.tif", filetype = "GTiff", overwrite = TRUE)
```

