---
title: "Substrate Raster Cover"
author: "Jessica French"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r label = "read in libraries}
# read in packages
library(terra)
library(tmap)
library(sf)
library(tidyverse)
```


The purpose of this is to check the files that have been combined and/ or converted to raster that they map correctly etc. 

```{r label = "set data directory"}
# Define data directory
data_dir <- "/capstone/kelpgeomod/google_drive_download"
```

## Read in and Reclassify

For each file the substrate classification is different. After reading in each file a reclassification is created where the first column is the current value and the second column is the new value. This was determined by looking at the classification of each file in the markdown it was created and by comparing maps of each file. 

```{r label = "read in CA_maps file"}
# read in data from intermediate data substrate folder
CA_maps <- rast(file.path(data_dir, "02-intermediate-data/08-substrate-intermediate/seafloormapping-islands-substrate.tif")) 


plot(CA_maps)
```

```{r label = "read in USGS_islands file"}
# read in data from intermediate data folder
USGS_islands <- rast(file.path(data_dir, "02-intermediate-data/08-substrate-intermediate/usgs-island-substrate.tif")) |> as.numeric()

# create reclassification matrix
# before reclassification 0 = hard, 1 = mixed, and 2 = soft

USGS_islands_rcl <- matrix(c(0, 1, 1, 2, 2, 0), nrow = 3, ncol = 2, byrow = T)

# after reclassification 0 = soft, 1 = hard, and 2 = mixed

USGS_islands <- classify(USGS_islands, rcl = USGS_islands_rcl)

plot(USGS_islands)
```

```{r label = "read in USGS_coast data form intermediate data folder"}
USGS_coast <- rast(file.path(data_dir, "02-intermediate-data/08-substrate-intermediate/usgs-coast-substrate.tif")) |> 
  as.numeric()

#create reclassification matrix
# before reclassification 0 = anthropogenic, 1 = hard, 2 = mixed, and 3 = soft
USGS_coast_rcl <- matrix(c(4, 0, 2, 1, 1, 3, 3, 2), nrow = 4, ncol = 2, byrow = T)

# after reclassification 0 = soft, 1 = hard, 2 = mixed, adn 3 = anthropogenic. 

USGS_coast <- classify(USGS_coast, rcl = USGS_coast_rcl)

plot(USGS_coast)
```

```{r labels = "create dataframe of factor levels"}
# create data frame of desired factor levels to feed into the levels so that all of the rasters are the same 

relevel_df <- tribble(
  ~ "value", ~ "substrate_class", 
  0, "soft", 
  1, "hard", 
  2, "mixed", 
  3, "anthropogenic"
) 
```


```{r label = "combine files"}
# Combine files using merge, this is a simpler operation but none of these files overlap at this point so it should not be an issue and there is no risk of it doing any calculations with categorical variables.
SBC_substrate <- merge(CA_maps, USGS_islands, USGS_coast)

levels(SBC_substrate) <- relevel_df


# plot
plot(SBC_substrate) 

```

```{r label = "write to GTiff"}
# An aux.xml metadata file is automatically written with each file. 
# Write to intermediate folder
terra::writeRaster(SBC_substrate, "/capstone/kelpgeomod/google_drive_download/02-intermediate-data/08-substrate-intermediate/combined-sbchannel-substrate.tif", filetype = "GTiff", overwrite = TRUE)

# Write to deliverable folder
terra::writeRaster(SBC_substrate, "/capstone/kelpgeomod/google_drive_download/04-deliverables/04-combined-sbchannel-substrate/combined-sbchannel-substrate.tif", filetype = "GTiff", overwrite = TRUE)
```


