---
title: "Arrange Quarterly Data"
author: "Elke Windschitl"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: html_document
---

This is a script that gathers oceanographic rasters from across our data sets, formats them for Wallace Maxent analysis, and saves them to easy-to-use folders.

```{r}
library(tidyverse)
library(raster)
library(terra)
```

# Load in data using terra, assign layer names, then convert to package Raster objects

## SST

```{r}
# create a list to store the raster objects
sst_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in sst quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/intermediate_data/sst/sst_quarter_", i, ".tif")
  # create the raster object and add it to the list
  sst_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(sst_list[[i]]) <- "sst"
  sst_list[[i]] <- raster(sst_list[[i]])
}
```

## Nitrate + Nitrite

```{r}
# create a list to store the raster objects
nitrate_nitrite_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in nitrate_nitrite quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/analysis_data/inteprolate_idw/nitrate_nitrite/nitrate_nitrite_", i, ".tif")
  # create the raster object and add it to the list
  nitrate_nitrite_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(nitrate_nitrite_list[[i]]) <- "nitrate_nitrite"
  nitrate_nitrite_list[[i]] <- raster(nitrate_nitrite_list[[i]])
}
```

## Phosphate

```{r}
# create a list to store the raster objects
phosphate_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in phosphate quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/analysis_data/inteprolate_idw/phosphate/phosphate_", i, ".tif")
  # create the raster object and add it to the list
  phosphate_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(phosphate_list[[i]]) <- "phosphate"
  phosphate_list[[i]] <- raster(phosphate_list[[i]])
}
```

## Ammonium -- This is the processing step for ammonium, however, we do not include it in maxent to to lack of data.

```{r}
# create a list to store the raster objects
ammonium_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in ammonium quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/analysis_data/inteprolate_idw/ammonium/ammonium_", i, ".tif")
  # create the raster object and add it to the list
  ammonium_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(ammonium_list[[i]]) <- "ammonium"
  ammonium_list[[i]] <- raster(ammonium_list[[i]])
}
```

## Depth -- Althought depth is static over all quarters, we made four copies for ease of inclusion with other variables.

```{r}
# create a list to store the raster objects
depth_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in ammonium quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/intermediate_data/depth/SBC_depth.tif")
  # create the raster object and add it to the list
  depth_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(depth_list[[i]]) <- "depth"
  depth_list[[i]] <- raster(depth_list[[i]])
}

```

# Mask data to only include cells where we have values for every oceanographic factor.


```{r}
# define a function that checks if all values are not NA
all_not_na <- function(sst, nitrate_nitrite, phosphate, depth) {
  ifelse(!is.na(sst) & !is.na(nitrate_nitrite) & !is.na(phosphate) & !is.na(depth), 1, NA)
}

# apply the function to the rasters using overlay
maxent_mask_1 <- raster::overlay(sst = sst_list[[1]], 
                         nitrate_nitrite = nitrate_nitrite_list[[1]], 
                         phosphate = phosphate_list[[1]], 
                         depth = depth_list[[1]],
                         fun = all_not_na)

# plot the resulting raster
plot(maxent_mask)

dep <- mask(dep_raster, maxent_mask)
nit <- mask(nit_raster, maxent_mask)
pho <- mask(pho_raster, maxent_mask)



# Read in the three rasters
raster1 <- sst_list[[1]]
raster2 <- nitrate_nitrite_list[[1]]
raster3 <- phosphate_list[[1]]
raster4 <- depth_list[[1]]

# define a function that checks if all values are not NA
all_not_na <- function(w, x, y, z) {
  ifelse(!is.na(w) & !is.na(x) & !is.na(y) & !is.na(z), 1, NA)
}

# apply the function to the rasters using overlay
maxent_mask <- overlay(raster1, raster2, raster3, raster4, fun = all_not_na)

# plot the resulting raster
plot(maxent_mask)

```

