---
title: "Arrange Quarterly Data"
author: "Elke Windschitl"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: html_document
---

This is a script that gathers oceanographic rasters from across our data sets, formats them for Wallace Maxent analysis, and saves them to easy-to-use folders.

```{r}
library(tidyverse)
library(raster)
library(terra)
library(sf)
```

# Load in data using terra, assign layer names, then convert to package Raster objects

## SST

```{r}
# create a list to store the raster objects
sst_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in sst quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/02-intermediate-data/sst/sst_quarter_", i, ".tif")
  # create the raster object and add it to the list
  sst_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(sst_list[[i]]) <- "sst"
  sst_list[[i]] <- raster(sst_list[[i]])
}
```

## Nitrate + Nitrite

```{r}
# create a list to store the raster objects
nitrate_nitrite_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in nitrate_nitrite quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/inteprolate_idw/nitrate_nitrite/nitrtate_nitrite_", i, ".tif")
  # create the raster object and add it to the list
  nitrate_nitrite_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(nitrate_nitrite_list[[i]]) <- "nitrate_nitrite"
  nitrate_nitrite_list[[i]] <- raster(nitrate_nitrite_list[[i]])
}
```

## Phosphate

```{r}
# create a list to store the raster objects
phosphate_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in phosphate quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/inteprolate_idw/phosphate/phosphate_", i, ".tif")
  # create the raster object and add it to the list
  phosphate_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(phosphate_list[[i]]) <- "phosphate"
  phosphate_list[[i]] <- raster(phosphate_list[[i]])
}
```

## Ammonium -- This is the processing step for ammonium, however, we do not include it in maxent to to lack of data.

```{r}
# create a list to store the raster objects
ammonium_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in ammonium quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/inteprolate_idw/ammonium/ammonium_", i, ".tif")
  # create the raster object and add it to the list
  ammonium_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(ammonium_list[[i]]) <- "ammonium"
  ammonium_list[[i]] <- raster(ammonium_list[[i]])
}
```

## Depth -- Althought depth is static over all quarters, we made four copies for ease of inclusion with other variables.

```{r}
# create a list to store the raster objects
depth_list <- list()

# loop through quarters 1-4
for (i in 1:4) {
  # Read in ammonium quarterly data --------------------
  # create the file name
  file_name <- paste0("/capstone/kelpgeomod/new_file_structure/02-intermediate-data/depth/depth.tif")
  # create the raster object and add it to the list
  depth_list[[i]] <- rast(file_name, lyr = 1)
  
  # Process to match maxent needs -----------------
  names(depth_list[[i]]) <- "depth"
  depth_list[[i]] <- raster(depth_list[[i]])
}

```

### Mask data to only include cells where we have values for every oceanographic factor.


```{r}
for (i in 1:4){
  # Read in the three rasters
  #sst <- sst_list[[i]]
  nitrate_nitrite <- nitrate_nitrite_list[[i]]
  phosphate <- phosphate_list[[i]]
  depth <- depth_list[[i]]
  
  # define a function that checks if all values are not NA
  all_not_na <- function(nitrate_nitrite, phosphate, depth) {
    ifelse(!is.na(nitrate_nitrite) & !is.na(phosphate) & !is.na(depth), 1, NA)
    }
  # apply the function to the rasters using overlay
  maxent_mask <- overlay(nitrate_nitrite, phosphate, depth, fun = all_not_na)
  assign(paste0("maxent_mask_", i), maxent_mask)
}
```

### Apply the mask to every raster and export

```{r}
for (i in 1:4) {
  # create the maxent_mask for this iteration
  maxent_mask <- get(paste0("maxent_mask_", i))
  
  # mask the input rasters with the maxent_mask
  depth_i <- mask(depth_list[[i]], maxent_mask)
  nitrate_nitrite_i <- mask(nitrate_nitrite_list[[i]], maxent_mask)
  phosphate_i <- mask(phosphate_list[[i]], maxent_mask)
  
  # save the masked rasters to files with unique names
  writeRaster(depth_i, paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_", i, "/depth_", i, ".tif"), overwrite = TRUE)
  
  writeRaster(nitrate_nitrite_i, paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_", i, "/nitrate_nitrite_", i, ".tif"), overwrite = TRUE)
  
  writeRaster(phosphate_i, paste0("/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_", i, "/phosphate_", i, ".tif"), overwrite = TRUE)
} 

```

### Add AOI to each file 

```{r}
# read in the shapefile of aoi
aoi_shp <- st_read("/capstone/kelpgeomod/new_file_structure/01-raw-data/New_AOI_SBchannel_shp/New_AOI_SBchannel.shp")

# write the aoi object to a shapefile in each quarter file
st_write(aoi_shp, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_1/aoi.shp", append = FALSE)
st_write(aoi_shp, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_2/aoi.shp", append = FALSE)
st_write(aoi_shp, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_3/aoi.shp", append = FALSE)
st_write(aoi_shp, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_4/aoi.shp", append = FALSE)


# my_aoi <- coordinates <- tribble(
#   ~latitude, ~longitude,
#   30.59, -123.32, 
#   30.59, -115.70, 
#   36.08, -123.32, 
#   36.08, -115.70
# ) %>% relocate(longitude, .before = "latitude")
# 
# write_csv(my_aoi, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_1/sbc_aoi.csv", 
#           append = FALSE)
# write_csv(my_aoi, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_2/sbc_aoi.csv", 
#           append = FALSE)
# write_csv(my_aoi, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_3/sbc_aoi.csv", 
#           append = FALSE)
# write_csv(my_aoi, "/capstone/kelpgeomod/new_file_structure/03-analysis-data/maxent/quarter_4/sbc_aoi.csv", 
#           append = FALSE)
```

