---
title: "Synthesize Dataset"
author: "Javier Patrón"
date: "`r Sys.Date()`"
output: html_document
---

Load the libraries
```{r}
library(tidyverse)
library(raster)
library(sf)
library(sp)
library(dplyr)
library(tidyr)
library(tmap)
library(lubridate)
```


Set the kelpGeoMod Google Drive Working Directory
```{r}
#---- Set up the data directory (wherever your download of our Google Shared Drive lives)
drive_dir <- "/capstone/kelpgeomod/google_drive_download"
```

# 1) Read in the data
Read all the data sets
```{r}
# NUTRIENTS
# Observed Nutrients
joined_nutrients <- read_csv(file.path(drive_dir, "02-intermediate-data/09-nutrients-intermediate/joined/joined-nutrients.csv")) %>% 
  st_as_sf(coords = c("lon", "lat"))

# Phosphate
interp_phosphate <- brick(file.path(drive_dir, "03-analysis-data/02-nutrient-interpolation-analysis/phosphate/phosphate-quarter-brick.tif"))
obs_phosphate <- brick(file.path(drive_dir, "02-intermediate-data/09-nutrients-intermediate/joined/phosphate-brick.tif"))

# Nitrate Nitrite
interp_nitrate_nitrite <- brick(file.path(drive_dir, "03-analysis-data/02-nutrient-interpolation-analysis/nitrate-nitrite/nitrate-nitrite-quarter-brick.tif"))
obs_nitrate_nitrite <- brick(file.path(drive_dir, "02-intermediate-data/09-nutrients-intermediate/joined/nitrate-nitrite-brick.tif"))


#Ammonium
interp_ammonium <- brick(file.path(drive_dir, "03-analysis-data/02-nutrient-interpolation-analysis/ammonium/ammonium-quarter-brick.tif"))
obs_ammonium <- brick(file.path(drive_dir, "02-intermediate-data/09-nutrients-intermediate/joined/ammonium-brick.tif"))


# SEA SURFACE TEMPERATURE
sst <- brick(file.path(drive_dir, "02-intermediate-data/07-sst-intermediate/sst-brick.tif"))

# DEPTH
depth <- raster(file.path(drive_dir, "02-intermediate-data/06-depth-intermediate/depth.tif"))

#KELP

# Kelp Area
kelp_area <- brick(file.path(drive_dir,"02-intermediate-data/05-kelp-area-biomass-intermediate/kelp-area-brick.tif"))

#Kelp Biomass
kelp_biomass <- brick(file.path(drive_dir,"02-intermediate-data/05-kelp-area-biomass-intermediate/kelp-biomass-brick.tif"))


# MASK
mask <- raster(file.path(drive_dir,"02-intermediate-data/03-mask-intermediate/main-mask.tif"))

```

# 2) Transforming raster data to data frames
## DEPTH
```{r}
# Convert the Raster brick to a data frame
depth_df <- as.data.frame(depth, xy = TRUE, na.rm = TRUE)

# Define the years and quarters to be added
years <- 2014:2022
quarters <- 1:4

# Create an empty data frame to store the results
depth_year_quarter_df <- data.frame()

# Loop over quarters and years
for (q in quarters) {
  for (y in years) {
    # Add a row to the data frame with the current year and quarter
    row <- cbind(depth_df, year = y, quarter = q)
    depth_year_quarter_df <- rbind(depth_year_quarter_df, row)
  }
}


# Create a data frame for the Time Series
depth_year_quarter_df <- depth_year_quarter_df %>% 
  rename(lon = x, lat = y, depth = exportImage) %>% 
  dplyr::select(year, quarter, lon, lat, depth) %>%
  mutate(year = as.character(year),
         quarter = as.character(quarter))
  
```

# SST
```{r}
#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(sst)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(sst) <- new_names

# TESTING
# Check the extent and resolution of the raster object
identical(extent(sst), extent(mask))
identical(res(sst), res(mask))
identical(crs(sst), crs(mask))
identical(origin(sst), origin(mask))

# Convert the raster brick to a data frame
sst_year_quarter_df <- as.data.frame(sst, xy = TRUE, na.rm = TRUE)
sst_year_quarter_df <- sst_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "l" for "layer"
               names_to = "quarter_year", # Name of the new column to be created for year_quarter
               values_to = "sst") %>%     # Name of the new column to be created for sst values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and lon
  dplyr::select(year, quarter, lon, lat, sst)  # Reorder the columns as needed

```


# KELP
## Kelp Area
```{r}
# TESTING
# Check the extent and resolution of the raster_kelp object
identical(extent(kelp_area), extent(mask))
identical(res(kelp_area), res(mask))
identical(crs(kelp_area), crs(mask))
identical(origin(kelp_area), origin(mask))

#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(kelp_area)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(kelp_area) <- new_names


#Create a Data Frame from this list
# Convert the Raster brick to a data frame
kelp_area_year_quarter_df <- as.data.frame(kelp_area, xy = TRUE, na.rm = TRUE)
kelp_area_year_quarter_df <- kelp_area_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "Q"
               names_to = "quarter_year", # Name of the new column to be created for quarter-year
               values_to = "kelp_area") %>%     # Name of the new column to be created for SST values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the quarter_year column into quarter and year columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the Q of the quarter values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  dplyr::select(year, quarter, lon ,lat, kelp_area)  # Reorder the columns as needed
```

## Kelp Biomass
```{r}
# Re-add names since they are lost when you read back in
years <- c(2014:2022)
quarters <- c(1:4)

year_quarter <- c()
for (i in seq_along(years)) {
  for (j in seq_along(quarters)) {
  year_quarter_new <- paste0("layer", years[i], "_", quarters[j])
  year_quarter <- append(year_quarter, year_quarter_new)
  }
}

names(kelp_biomass) <- paste(year_quarter)

# TESTING
# Check the extent and resolution of the raster object
identical(extent(kelp_biomass), extent(mask))
identical(res(kelp_biomass), res(mask))
identical(crs(kelp_biomass), crs(mask))
identical(origin(kelp_biomass), origin(mask))

# Convert the raster brick to a data frame
kelp_biomass_year_quarter_df <- as.data.frame(kelp_biomass, xy = TRUE, na.rm = TRUE)
kelp_biomass_year_quarter_df <- kelp_biomass_year_quarter_df %>%
  pivot_longer(cols = starts_with("l"), # Select all columns that start with "l" for "layer"
               names_to = "year_quarter", # Name of the new column to be created for year_quarter
               values_to = "kelp_biomass") %>%     # Name of the new column to be created for kelp values
  separate(year_quarter, c("year", "quarter")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(year = str_remove(year, "layer")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  dplyr::select(year, quarter, lon, lat, kelp_biomass)  # Reorder the columns as needed

```

# NUTRIENTS
## Observed Nutrients
```{r}
# Extract the coordinates from the geometry column
coordinates <- st_coordinates(joined_nutrients)

# Convert the coordinates to a data frame
coordinates_df <- as.data.frame(coordinates)

# Rename the columns to lon and lat
names(coordinates_df) <- c("lon", "lat")

# Add the lon and lat columns to the original sf object
joined_nutrients_df <- cbind(joined_nutrients, coordinates_df) %>% 
  st_drop_geometry(joined_nutrients, "geometry")

```

## Observed Phosphate
```{r}
#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(obs_phosphate)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(obs_phosphate) <- new_names

# TESTING
# Check the extent and resolution of the raster object
identical(extent(obs_phosphate), extent(mask))
identical(res(obs_phosphate), res(mask))
identical(crs(obs_phosphate), crs(mask))
identical(origin(obs_phosphate), origin(mask))

# Convert the raster brick to a data frame
phosphate_year_quarter_df <- as.data.frame(obs_phosphate, xy = TRUE, na.rm = FALSE)
phosphate_year_quarter_df <- phosphate_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "l" for "layer"
               names_to = "quarter_year", # Name of the new column to be created for year_quarter
               values_to = "phosphate") %>%     # Name of the new column to be created for sst values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and lon
  dplyr::select(year, quarter, lon, lat, phosphate)  # Reorder the columns as needed
```

## Observed Nitrate Nitrite
```{r}
#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(obs_nitrate_nitrite)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(obs_nitrate_nitrite) <- new_names

#TESTING
# Check the extent and resolution of the raster object
identical(extent(obs_nitrate_nitrite), extent(mask))
identical(res(obs_nitrate_nitrite), res(mask))
identical(crs(obs_nitrate_nitrite), crs(mask))
identical(origin(obs_nitrate_nitrite), origin(mask))

# Convert the raster brick to a data frame
nitrate_nitrite_year_quarter_df <- as.data.frame(obs_nitrate_nitrite, xy = TRUE, na.rm = FALSE)
nitrate_nitrite_year_quarter_df <- nitrate_nitrite_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "l" for "layer"
               names_to = "quarter_year", # Name of the new column to be created for year_quarter
               values_to = "nitrate_nitrite") %>%     # Name of the new column to be created for sst values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and lon
  dplyr::select(year, quarter, lon, lat, nitrate_nitrite)  # Reorder the columns as needed
```

## Observed Ammonium
```{r}
#Create the names for the raster files
# Get current names of layers in raster_brick
current_names <- names(obs_ammonium)
# Define quarter abbreviations
quarter_name <- c("Q1", "Q2", "Q3", "Q4")
# Create vector of new names in desired format
new_names <- paste0(quarter_name, "-", rep(2014:2022, each = 4))
# Rename layers in raster_brick with new names
names(obs_ammonium) <- new_names

# TESTING
# Check the extent and resolution of the raster object
identical(extent(obs_ammonium), extent(mask))
identical(res(obs_ammonium), res(mask))
identical(crs(obs_ammonium), crs(mask))
identical(origin(obs_ammonium), origin(mask))

# Convert the raster brick to a data frame
ammonium_year_quarter_df <- as.data.frame(obs_ammonium, xy = TRUE, na.rm = FALSE)
ammonium_year_quarter_df <- ammonium_year_quarter_df %>%
  pivot_longer(cols = starts_with("Q"), # Select all columns that start with "l" for "layer"
               names_to = "quarter_year", # Name of the new column to be created for year_quarter
               values_to = "ammonium") %>%     # Name of the new column to be created for sst values
  separate(quarter_year, c("quarter", "year")) %>% # Separate the year_quarter column into year and quarter columns
  mutate(quarter = str_remove(quarter, "Q")) %>% # Removes the "layer" of the year values
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and lon
  dplyr::select(year, quarter, lon, lat, ammonium)  # Reorder the columns as needed

```

# Synthesized Data Set
## All values
```{r}

# Merge the all normalized data sets
all_sythesized_merge <- depth_year_quarter_df %>% 
  left_join(sst_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  left_join(kelp_area_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  left_join(kelp_biomass_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  left_join(nitrate_nitrite_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  left_join(phosphate_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  left_join(ammonium_year_quarter_df, by = c("year", "quarter", "lat", "lon")) %>% 
  relocate(lat, .before = lon)


# Upload the synthesize data to the organizded kelpGeoMod Google Drive 
write_csv(all_sythesized_merge, path = "/capstone/kelpgeomod/google_drive_download/04-deliverables/02-full-synthesized/full-synthesized.csv")

write_csv(all_sythesized_merge, path = "/capstone/kelpgeomod/google_drive_download/03-analysis-data/03-data-synthesization-analysis/full-synthesized.csv")

```

## Observed Synthesized Data Set

TESTING
Understand how many observed data points we have per year and quarter
```{r}
points <- joined_nutrients %>% 
  st_drop_geometry() %>% 
  group_by(year, quarter) %>% 
  summarise(count = n())

# Its important to notice that for the Year 2020 Q2, there are no points for the observed nutrients
```

Extract the raster values of depth, sst, kelp area, and kelp biomass with the observed nutrient data Point
```{r}

# Replicate the depth layer to create a raster brick with 35 layers we will extract the raster of 2020, quarter two. alias the 26 out of the 36 raster bricks

depth_brick <- stack(rep(list(depth), 35))

# Create an example raster list
raster_list <- list(kelp_biomass[[-26]], kelp_area[[-26]], sst[[-26]], depth_brick)

years <- rep(2014:2022, each = 4)[-26]
quarters <- rep(sequence(4), times = 9)[-26]

# Create an empty list to store extracted values
extracted_biomass_values <- list()
extracted_area_values <- list()
extracted_sst_values <- list()
extracted_depth_values <- list()


# Iterate over the years and quarters
for (i in 1:length(years)) {
  y <- years[i]
  q <- quarters[i]
  
  # Filter the joined_nutrients dataframe for the specific quarter
  filtered_nutrients <- subset(joined_nutrients, year == y & quarter == q)
  
  # Biomass
  biomass_values <- raster::extract(x = kelp_biomass[[i]], y = filtered_nutrients)
  extracted_biomass_values[[i]] <- biomass_values
  
  # Area
  area_values <- raster::extract(x = kelp_area[[i]], y = filtered_nutrients)
  extracted_area_values[[i]] <- area_values
  
  # SST
  sst_values <- raster::extract(x = sst[[i]], y = filtered_nutrients)
  extracted_sst_values[[i]] <- sst_values
  
  # Depth
  depth_values <- raster::extract(x = depth_brick[[i]], y = filtered_nutrients)
  extracted_depth_values[[i]] <- depth_values
 
}

```


```{r}
# Convert the individual exracted values lists to a data frame
obs_nutrients_sythesized <- data.frame(
  kelp_biomass = unlist(extracted_biomass_values),
  kelp_area = unlist(extracted_area_values),
  sst = unlist(extracted_sst_values),
  depth = unlist(extracted_depth_values)
) %>% 
  cbind(joined_nutrients_df) %>% 
  rename(nutrient_source = source) %>% 
  dplyr::select(year, quarter,lat, lon, temp, nitrate, nitrite, nitrate_nitrite, phosphate, ammonium, sst, nutrient_source, depth, kelp_area, kelp_biomass)

# Upload the synthesize data to our organized kelpGeoMod Google Drive
write_csv(obs_nutrients_sythesized, path = "/capstone/kelpgeomod/google_drive_download/04-deliverables/03-observed-nutrients-synthesized/observed-nutrients-synthesized.csv")

write_csv(obs_nutrients_sythesized, path = "/capstone/kelpgeomod/google_drive_download/03-analysis-data/03-data-synthesization-analysis/observed-nutrients-synthesized.csv")

```

TESTING
```{r}
# Checking the sythesized data
one <- skimr::skim(obs_nutrients_sythesized)

# Checking the depth data

depth_check <- list()
depth_values <- raster::extract(x = depth, y = joined_nutrients)
depth_check <- depth_values

check_df <- data.frame(depth = unlist(depth_check)
) %>% 
  cbind(joined_nutrients) %>%
  filter(is.na(depth)) %>% 
  group_by(geometry) %>% 
  summarise(depth = sum(depth))

# Convert the data frame to an sf object
check_df <- st_as_sf(check_df)
```

TESTING
Understand where are the point data and how it extracts the information for each raster layer
```{r}
# Display the tmap
tmap_mode("view")

# Create a tmap object
tm_shape(kelp_biomass[[20]]) + # Q4, 2018
  tm_raster(palette = "Greens") +
tm_shape(depth) + 
  tm_raster() +
tm_shape(joined_nutrients) +
  tm_symbols(col = "red", alpha = 0.8) +
tm_shape(check_df) +
  tm_symbols(col = "blue", size = 0.1)
```

# Time Series Interpolated Nutrients

## Interpolated Nutrients

### Phosphate
```{r}

#TESTING
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_phosphate), extent(mask))
identical(res(interp_phosphate), res(mask))
identical(crs(interp_phosphate), crs(mask))
identical(origin(interp_phosphate), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
phosphate_df <- as.data.frame(interp_phosphate, xy = TRUE, na.rm = TRUE)

names(phosphate_df)[3:6] <- c("1", "2", "3", "4")

phosphate_seasonal <- phosphate_df %>%
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
  pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "phosphate") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(phosphate = mean(phosphate, na.rm = T))
  
  
```

### Nitrate_Nitrite
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_nitrate_nitrite), extent(mask))
identical(res(interp_nitrate_nitrite), res(mask))
identical(crs(interp_nitrate_nitrite), crs(mask))
identical(origin(interp_nitrate_nitrite), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
nitrate_nitrite_df <- as.data.frame(interp_nitrate_nitrite, xy = TRUE, na.rm = TRUE)
nitrate_nitrite_seasonal <- nitrate_nitrite_df %>%
  rename(lon = "x", lat = "y") %>%  # Rename the x and y columns to lat and long
pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "nitrate_nitrite") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(nitrate_nitrite = mean(nitrate_nitrite, na.rm = T))
```

### Ammonium
```{r}
# Check the extent and resolution of the raster_kelp object
identical(extent(interp_ammonium), extent(mask))
identical(res(interp_ammonium), res(mask))
identical(crs(interp_ammonium), crs(mask))
identical(origin(interp_ammonium), origin(mask))

#Create a Data Frame from this list
# Convert the Raster brick to a data frame
ammonium_df <- as.data.frame(interp_ammonium, xy = TRUE, na.rm = TRUE)
ammonium_seasonal <- ammonium_df %>%
  rename(lon = "x", lat = "y")  %>%  # Rename the x and y columns to lat and long
  pivot_longer(cols = 3:6,
               names_to = "quarter",
               values_to = "ammonium") %>% 
  relocate(quarter, .before = lon) %>% 
  group_by(quarter, lon, lat) %>% 
  summarise(ammonium = mean(ammonium, na.rm = T))

```

# Time Series Visualization
## Normalize data sets
```{r}
# Compute time series wrangling for each year_quarter data frames
#DEPTH
depth_time_series <- depth_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(depth = mean(depth)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# SST
sst_time_series <- sst_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# KELP
# Kelp Area
kelp_area_time_series <- kelp_area_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(kelp_area = sum(kelp_area, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# Kelp Biomass
kelp_biomass_time_series <- kelp_biomass_year_quarter_df %>% 
  group_by(year, quarter) %>% 
  summarise(kelp_biomass = sum(kelp_biomass, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

# NUTRIENTS
nutrients_time_series <- joined_nutrients_df %>% 
  group_by(year, quarter) %>% 
  summarise(temp = mean(temp, na.rm = T),
            nitrate = mean(nitrate, na.rm = T),
            nitrite = mean(nitrite, na.rm = T),
            nitrate_nitrite = mean(nitrate_nitrite, na.rm = T),
            ammonium = mean(ammonium, na.rm = T),
            phosphate = mean(phosphate, na.rm = T)) %>% 
  mutate(year = as.factor(year)) %>% 
  mutate(quarter = as.factor(quarter))

```

## Join the time series data frames
```{r}
# Merge the all normalized data sets
time_series_merge <- depth_time_series %>% 
  left_join(sst_time_series, by = c("year", "quarter")) %>% 
  left_join(kelp_area_time_series, by = c("year", "quarter")) %>% 
  left_join(kelp_biomass_time_series, by = c("year", "quarter")) %>% 
  left_join(nutrients_time_series, by = c("year", "quarter"))

# Pivot the data into long format
time_series_long <- pivot_longer(time_series_merge, 
                                 cols = 3:11, 
                                 names_to = "variable", 
                                 values_to = "value") %>% 
   mutate(date = as.Date(paste(year, as.integer(quarter)*3, "01", sep = "-")))

# Upload the data to Tsosie
write_csv(time_series_merge, 
            path = "/capstone/kelpgeomod/google_drive_download/03-analysis-data/03-data-synthesization-analysis/time-series-synthesized.csv")
```

## Set the variable colors
```{r}
my_colors <- c("depth" = "#2E2E2E", 
               "sst" = "#FD8D3C", 
               "area" = "#3CB44B", 
               "biomass" = "#91CF60", 
               "temp" = "#984EF1", 
               "nitrate" = "#1F78B4", 
               "nitrite" = "#A6CEE3",
               "nitrate_nitrite" = "#984EA3", 
               "ammonium" = "#E41A1C", 
               "phosphate" = "#FF7F00")

```

## Facet Wrap Visualization
```{r}
ggplot(time_series_long, aes(x = date, y = value, color = variable)) +
  geom_line() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_color_manual(values = my_colors) +
  labs(x = "Date", y = "Value", color = "Variable", title = "Oceanographic Time Series") +
  facet_wrap(~ variable, scales = "free_y") +
  theme_minimal()

```

## Individual Visualization
```{r}

# Define the color palette
my_colors <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02", "#a6761d", "#666666", "#bdbdbd")

# Create a list of unique variable names
variable_names <- unique(time_series_long$variable)

# Create a for loop to create a separate ggplot for each variable
for (i in seq_along(variable_names)) {
  # Subset the data for the current variable
  data_subset <- subset(time_series_long, variable == variable_names[i])
  
  # Create a ggplot for the current variable
  p <- ggplot(data_subset, aes(x = date, y = value, color = variable_names[i])) +
    geom_line(color = my_colors[i]) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    labs(x = "Date", 
         y = variable_names[i], 
         title = "kelpGeoMod Time Series",
         subtitle = paste("Oceanographic factor:", variable_names[i])) +
    theme_minimal()
  
  # Print the ggplot
  print(p)
}

```

# Seasonal Rasters
## Normalize data sets
```{r}
# Compute wrangling for each data frame to get the seasonal mean
#DEPTH
depth_seasonal <- depth_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(depth = mean(depth)) %>% 
  mutate(quarter = as.factor(quarter))

#SST
sst_seasonal <- sst_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(quarter = as.factor(quarter))

#KELP
# Kelp Area
kelp_area_seasonal <- kelp_area_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(kelp_area = mean(kelp_area, na.rm = T)) %>% 
  mutate(quarter = as.factor(quarter))

# Kelp Biomass
kelp_biomass_seasonal <- kelp_biomass_year_quarter_df %>% 
  group_by(quarter, lat, lon) %>% 
  summarise(kelp_biomass = mean(kelp_biomass, na.rm = T)) %>% 
  mutate(quarter = as.factor(quarter))

```

## Seasonal Join
```{r}
# Merge the all normalized seasonal data sets
seasonal_merge <- depth_seasonal %>% 
  left_join(sst_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(kelp_area_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(kelp_biomass_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(phosphate_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(nitrate_nitrite_seasonal, by = c("quarter", "lon", "lat")) %>% 
  left_join(ammonium_seasonal, by = c("quarter", "lon", "lat"))
  
```

Save the 36 raster brick
```{r}
# Write RasterBrick Quarter object to file
write_csv(seasonal_merge, "/capstone/kelpgeomod/google_drive_download/03-analysis-data/03-data-synthesization-analysis/quarterly-interpolated-synthesized.csv")
```





